Complete draft

### 2019-05-27 doc/point.tex
added appendix deriving estimators.  Need to check minus sign

The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,

@@

The Gaussian likelihood as a function of $\theta$
\[
\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
\]\par
To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
\begin{equation}
-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
\label{log:eqn}
\end{equation}

The estimator and Fisher Matrix for this Gaussian likelihood are known
\begin{equation}
\hat{\theta}_\alpha = \frac{1}{2} F^{-1}_{\alpha \beta} [(\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]
\end{equation}
and
\begin{equation}
F_{\alpha\beta} = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0 \textbf{S}_\beta).
\end{equation}

The term
\begin{align}
\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \textbf{x}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \textbf{x}\\
&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
\end{align*}

@@

First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
\begin{equation}
\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M}.
\end{equation}
Expanded as a Taylor series
\begin{align}
\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
& = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
\end{align}
%= -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
%\] \par

A term of interest is 
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
 $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
and $\xi_{vv}$.  In this manner
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}
\begin{comment}

Now consider the next term $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}$.  The middle section is
\begin{align}
\textbf{S}\textbf{C}_0^{-1}\textbf{S} & =
\begin{pmatrix}
 \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
 \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
\end{pmatrix}
\end{align}
to leading order of $\bar{N}$ anticipating that we will take the limit where it goes to zero.
Then
\begin{align}
\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}
& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
 + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
\end{align}

We have assembled the contributions of $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ to second order in $A$.

Returning to the other contribution to the log-likelihood in Eq.~\ref{log:eqn} we have
\begin{align*}
\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\textbf{X})\textbf{M})) \\
&= \ln\det(\textbf{I}+A\textbf{X}) +\text{const} \\
&= A \tr \textbf{X} -\frac{A^2}{2}\tr \textbf{X}^2+ O(A^3) + \ldots\\
\end{align*}
Now
\begin{equation}
\textbf{X} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
\begin{pmatrix}
\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
\end{pmatrix}.
\end{equation}
The contribution to the trace of the upper diagonal in the limit of small cell volumes is
\begin{equation}
\bar{N} \tr \xi_{\delta\delta} = n \int \xi_{\delta\delta,VV} dV
\end{equation}
and the lower diagonal is
\begin{equation}
\tr  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}}   \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}.
\end{equation}

The trace of $\textbf{X}^2$ has terms, to leading order in $\bar{N}$, has terms
\begin{equation}
\bar{N}  \tr  \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} = n \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV
\end{equation}
and 
\begin{equation}
\tr   \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}} \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}
\end{equation}

@@

\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M},

@@

\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.

@@

s &=\textbf{x}^T \textbf{C}^{-1}\textbf{x} \\
&= -A\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right).

>>>>

%The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected, 
The data themselves are denoted as $\textbf{y} = [N_1, N_2, \hdots, N_{N_{cell}}, v_1, \hdots, v_{N_{gal}}]$
and are modeled as $\textbf{y}_0 =[\bar{N}, \bar{N}, \hdots, \bar{N}, v^0_{1}, \hdots, v^0_{N_{gal}}]$.  For now we assume that the model is not fixed but depends
on linear parameters $p_\alpha$ such that
 $\textbf{y}_0 = p_\alpha \textbf{x}_\alpha$.
Their difference is  denoted as $\Delta \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\Delta v= v-v^0$.  (Although the note is written in terms of velocities
in practice we expect to work with magnitudes as the are observed.)

@@

The Gaussian likelihood as a function of $\theta$
\[
\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}\right\}.
\]\par
To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
\begin{equation}
-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y} + \text{const}.
\label{log:eqn}
\end{equation}

The estimator and Fisher matrix for this Gaussian likelihood are
\begin{align}
\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]\\
\hat{p}_\alpha  &=  F^{(pp)\, -1}_{\alpha \beta}  (\textbf{y}^T  \textbf{C}^{-1} x_\beta)
\end{align}
where  
\begin{align}
F^{(\theta\theta)}_{\alpha\beta}& = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1}  \textbf{S}_\beta)\\
F^{(pp)}_{\alpha\beta} & =x^T_\alpha \textbf{C}^{-1} x_\beta\\
F^{(\theta p)}_{\alpha\beta} & = 0.
\end{align}

We focus on the parameters $\theta$ related to the covariance matrix .
The term
\begin{align}
\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \Delta \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \Delta \textbf{y}\\
&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
\begin{align*}
\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
\end{align*}

@@

First, we focus on the  term $\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
by   introducing the matrices $\textbf{M}$ and $\Delta \textbf{y}$,
where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \Delta \textbf{y} \textbf{M}$ such that $\Delta \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\Delta \textbf{y}$ and $\textbf{M}$:
\begin{equation}
\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M}.
\end{equation}
Expanded as a Taylor series
\begin{align}
\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\Delta \textbf{y}+A^2 \Delta \textbf{y}^2 + \ldots) \textbf{M}^{-1} \\
& = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
\end{align}
%= -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
%\] \par

A term of interest is 
\begin{align}
\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

We evaluate $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ by splitting up the counts and velocities.  The data vector $\Delta \textbf{y}$ is composed of the counts
 $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
and $\xi_{vv}$.  In this manner
\begin{align}
\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
\end{align}
Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
\begin{align*}
\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
\end{align*}

Now consider the next term $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}$.  The middle section is
\begin{align}
\textbf{S}\textbf{C}_0^{-1}\textbf{S} & =
\begin{pmatrix}
 \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
 \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
\end{pmatrix}
\end{align}
to leading order of $\bar{N}$ anticipating that we will take the limit where it goes to zero.
Then
\begin{align}
\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}
& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
 + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
\end{align}

We have assembled the contributions of $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ to second order in $A$.

Returning to the other contribution to the log-likelihood in Eq.~\ref{log:eqn} we have
\begin{align*}
\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M})) \\
&= \ln\det(\textbf{I}+A\Delta \textbf{y}) +\text{const} \\
&= A \tr \Delta \textbf{y} -\frac{A^2}{2}\tr \Delta \textbf{y}^2+ O(A^3) + \ldots\\
\end{align*}
Now
\begin{equation}
\Delta \textbf{y} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
\begin{pmatrix}
\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
\end{pmatrix}.
\end{equation}
The contribution to the trace of the upper diagonal in the limit of small cell volumes is
\begin{equation}
\bar{N} \tr \xi_{\delta\delta} = n \int \xi_{\delta\delta,VV} dV
\end{equation}
and the lower diagonal is
\begin{equation}
\tr  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}}   \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}.
\end{equation}

The trace of $\Delta \textbf{y}^2$ has terms, to leading order in $\bar{N}$, has terms
\begin{equation}
\bar{N}  \tr  \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} = n \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV
\end{equation}
and 
\begin{equation}
\tr   \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}} \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}
\end{equation}

@@

\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M},

@@

\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\Delta \textbf{y}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.

@@

s &=\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y} \\
&= -A\left(\Delta \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\Delta \textbf{y}\right).
